---
title: Who uses Socrata?
tweet_text: When I downloaded all of the @socrata datasets, I also downloaded all of the users.
---

(Introduction here)

## The user data
I'm still using the same 80,000 JSON files that I downloaded last month.
Each of these metadata files describes a Socrata view. Among their many
fields are two fields related to the people who interacted with that view.
These fields are the `owner` and  `tableAuthor` fields. These fields are
both hashes/dictionaries/arrays with user information. For example, here's
mine.

    {
        "id" : "x73b-9d8f",
        "displayName" : "Thomas Levine",
        "emailUnsubscribed" : false,
        "privacyControl" : "login",
        "profileLastModified" : 1371141485,
        "screenName" : "Thomas Levine"
    }

### `owner`
The `owner` field is quite straightforward; it's simply the person who
created the view, be it a dataset, chart, map or whatever.

### `tableAuthor`
The `tableAuthor` field is the person who uploaded the dataset associated
with the view. For datasets, this is the same as the owner, but for charts,
maps and filtered views, it might be different; one person may have uploaded
the data, and another person might have made the chart. "view", "dataset",
"map", "chart", "filtered view" and "table" all have rather special meanings
inside Socrata, so you should read my [post on that](/!/socrata-genealogies)
if the previous sentence didn't make sense.

## Extracting the user data
I already had a system for storing all of the Socrata metadata, so I just needed to write a
[sloppy function](https://github.com/tlevine/socrata-analysis/blob/e3d12254e928f986f1a83ed2099577e289048b94/numbers/run.py#L207)
to query it. This resulted in an SQLite3 database, which I converted to a
[CSV file](https://raw.github.com/tlevine/socrata-analysis/e3d12254e928f986f1a83ed2099577e289048b94/users.csv)
and read into R.

<!--
In case a view was federated, only one copy is counted, so this doesn't
have the duplication issues that some people were concerned about.
-->

Rather than remembering which tables and datasets everyone owned, that function
just remembers how many. Maybe I'll do something with the links later.

## Analysis
The table schema looks like this.

```{r schema}
str(users)
```

### Missing data
Not everyone had all of the fields. For example, hardly anyone has a profile image.

```{r profile.image}
table(is.na(users$profileImageUrlLarge))
```

```{r missingness}
a <- t(as.data.frame(lapply(users, function(variable){ sum(is.na(variable)) } )))
missingness <- data.frame(
  variable = rownames(a),
  present.value = a[,1]
)
missingness <- missingness[order(missingness$present.value, decreasing = T),]
barplot(
  (missingness$present.value),
  names.arg = missingness$variable,
  border = NA, col = 'grey', axes = F, horiz = T, las = 1,
  xlim = c(0, 12000),
  main = 'How complete is the user profile metadata?',
  xlab = 'Number of users with the field',
  ylab = 'Field'
)
axis(1, at = seq(0, 12000, 2000))
abline(v = nrow(users), col = 2, lty = 2)
text(x = nrow(users), y = 15, labels = paste(nrow(users), 'total users'), col = 2, pos = 2)
text(x = 2500, y = 14, labels = 'Missing data', col = 2)
text(x = 10000, y = 3, labels = 'Complete data', col = 2)
```

### Summary statistics
Let's start by looking at those variables for which no data are missing.
