The first of Sunlight Foundation's 32
[Open Data Policy Guidelines](http://sunlightfoundation.com/opendataguidelines/)
is to "Set The Default To Open".

> Most public records systems, including the Freedom of Information Act itself, are systems of reactive disclosure -- meaning that a question has to be asked before an answer given; public information requested, before it is disclosed.
>
> Proactive disclosure is the opposite. Proactive disclosure is the release of public information -- online and in open formats (see Provisions 8 and 9) -- before it is asked for. This is no simple task, but, in a way, it’s what all "open data" is aiming to accomplish. Setting the default to open means that the government and parties acting on its behalf will make public information available proactively and that they’ll put that information within reach of the public (online), with low to no barriers for its reuse and consumption. Open formats may help us maximize on the value we can extract from certain kinds of public data today, but to ensure that data publishing is sustained and, in fact, made easier over time, we need to reset the default for how data is released and disclosed.
>
> Setting the default to open is about living up to the potential of our information, about looking at comprehensive information management, and making determinations that fall in the public interest. It’s about purely practical government improvements, too, and taking steps that not only keep government systems up to date, but ensure that we have the foresight to survive changes in technology that we can’t predict.
>
> Usually, for information to be defined as public, important restrictions have already been applied. Therefore, policy language can be used to outline that "all public data and information must be considered open and accessible." Whether listed as part of a statement of intent (as Austin, Texas does; a concept explored more in Provision 21), as direction to a new oversight authority (Provision 22), or as the underlying aim of new data guidance (Provision 20), openness by default is a critical tool in crafting open data policies that are both ambitious and sustainable.

Software can encourage this practice of making data open by default.

## Display types on Socrata portals
I previously [downloaded](/!/socrata-summary) metadata about all of
the datasets on all of the Socrata portals, and I continue to find
interesting things in these data.

Let's focus on the things that are neither tables nor external links.

```{r not_boring}
not.boring <- subset(socrata.distinct, displayType != '' & displayType != 'table' & displayType != 'href' & displayType != 'blob')
not.boring$displayType <- factor(not.boring$displayType, levels = names(sort(table(not.boring$displayType), decreasing = T)))
ggplot(not.boring) + aes(x = displayType) + geom_bar() + coord_flip()
```

I was somewhat surprised to see forms and calendars in the portals.
I've [previously](/!/open-calendars) written about why I think Socrata calendars are cool,
so now I'm just going to talk about forms.

## Forms
Much of the goal of these portals is to open up existing government data, but
[forms](https://data.wa.gov/Economics/Broadband-Project-Data-Entry/38rz-krmg?) provide a way for citizens to create data.
lets you enter data. A bunch of people have implemented them, but none seems to get accessed much.

```{r form_use_3}
forms <- sqldf('select portal, sum(viewCount) AS "total.view.count", count(*) AS "number.of.forms" FROM socrata where displayType = "form" group by portal')
ggplot(forms) +
  aes(x = number.of.forms, y = total.view.count, label = portal) + geom_text() +
```
I'm gonna remove opendata.socrata.com to make that easier to read.

```{r form_use_4}
forms <- sqldf('select portal, sum(viewCount) AS "total.view.count", count(*) AS "number.of.forms" FROM socrata where displayType = "form" and portal != "opendata.socrata.com" group by portal')
ggplot(forms) +
  aes(x = number.of.forms, y = total.view.count, label = portal) + geom_text() +
  scale_x_continuous('Number of forms in the portal') +
  scale_y_continuous('Total hits across all forms', labels = comma)
```

### Feedback
I hadn't seen [nmfs.socrata.com](https://nmfs.socrata.com) before.
It belongs to the [National Oceanic and Atmospheric Administration Fisheries Service](http://www.nmfs.noaa.gov),
which apparently used [a Socrata form](https://nmfs.socrata.com/Government/2011-Aquaculture-Public-Comments-Form/u5id-8nqp) to power a
[policy comments website](http://www.nmfs.noaa.gov/aquaculture/policy2/).

New York made a form for [feedback on the portal](https://data.ny.gov/dataset/Give-Feedback/fq3e-q75i?).

### Calendar
World Bank Open Finances made a
[form](https://finances.worldbank.org/dataset/Global-Open-Data-Calendar-Entry-Form/qdbh-rfd3?)
that populates an
[open data events calendar](https://finances.worldbank.org/dataset/Global-Open-Data-Calendar/g4sx-dwxc).

## Relevance to software
I see data portal software as sending data from publisher silos to an open, magical website.
How about about prevending data from being siloed in the first place? I previously
[hinted](http://thomaslevine.com/!/socrata-calendars#opening-data-at-their-sources) at this,
but now I have some clear ideas.

The three examples of Socrata forms show us how we can turn user input on a website into
open data automatically. Using a Socrata form to compose a dataset is quite inconvenient,
unreliable, limited, and other bad things, but we can apply this concept with other software.

### Opening user-entered data on your website
If you have a website that stores information in a standard database (like MySQL) and
you separate the private information from the public information, you can quite safely
and easily have it sent to a data portal.

### Storing application data directly in a data portal
If you have a simple website, maybe you don't have to run your own database
and write your own web APIs. You could store the data directly in the data portal
and query it from the data portal. If this is powerful enough for you, it
simplifies both your database management and your opening of data.

### Opening data from some other web service
If you want to avoid managing your own database and web APIs,
you'll probably find
[some](http://keen.io)
[other](https://www.parse.com)
[service](http://stormpath.com/)
[that](http://www.google.com/analytics/)
[does](www.iriscouch.com)
[this](https://scraperwiki.com/)
better than data portal software does.
And even if you aren't running something that you think of as a "website",
you are probably storing some sort of data in
[some](https://www.google.com/calendar)
[web](http://www.salesforce.com/)
[service](http://mailchimp.com)

But these other services tend not to focus on cataloging disparate datasets,
converting them to different formats, &c. Said otherwise, they don't try to
open the data. It's not like they're particularly trying to keep the data
closed either; it's probably more that their customers haven't asked them for
open data features.

When a lot of people use services like these, the services' protocals
naturally become standard, so it becomes worthwhile to write tools that pull
data from these services into some standard place like a data portal.

Let's hope that something like [dat](https://github.com/maxogden/dat) 
