The first of Sunlight Foundation's 32
[Open Data Policy Guidelines](http://sunlightfoundation.com/opendataguidelines/)
is to "Set The Default To Open".

> Most public records systems, including the Freedom of Information Act itself, are systems of reactive disclosure -- meaning that a question has to be asked before an answer given; public information requested, before it is disclosed.
>
> Proactive disclosure is the opposite. Proactive disclosure is the release of public information -- online and in open formats (see Provisions 8 and 9) -- before it is asked for. This is no simple task, but, in a way, it’s what all "open data" is aiming to accomplish. Setting the default to open means that the government and parties acting on its behalf will make public information available proactively and that they’ll put that information within reach of the public (online), with low to no barriers for its reuse and consumption. Open formats may help us maximize on the value we can extract from certain kinds of public data today, but to ensure that data publishing is sustained and, in fact, made easier over time, we need to reset the default for how data is released and disclosed.
>
> Setting the default to open is about living up to the potential of our information, about looking at comprehensive information management, and making determinations that fall in the public interest. It’s about purely practical government improvements, too, and taking steps that not only keep government systems up to date, but ensure that we have the foresight to survive changes in technology that we can’t predict.
>
> Usually, for information to be defined as public, important restrictions have already been applied. Therefore, policy language can be used to outline that "all public data and information must be considered open and accessible." Whether listed as part of a statement of intent (as Austin, Texas does; a concept explored more in Provision 21), as direction to a new oversight authority (Provision 22), or as the underlying aim of new data guidance (Provision 20), openness by default is a critical tool in crafting open data policies that are both ambitious and sustainable.

Software can encourage this practice of making data open by default.

## Display types on Socrata portals
I previously [downloaded](/!/socrata-summary) metadata about all of
the datasets on all of the Socrata portals, and I continue to find
interesting things in these data.

Let's focus on the things that are neither tables nor external links.

```{r not_boring}
not.boring <- subset(socrata.distinct, displayType != '' & displayType != 'table' & displayType != 'href' & displayType != 'blob')
not.boring$displayType <- factor(not.boring$displayType, levels = names(sort(table(not.boring$displayType), decreasing = T)))
ggplot(not.boring) + aes(x = displayType) + geom_bar() + coord_flip()
```

## Surprising display types
I was somewhat surprised to see forms and calendars in the portals.

### Form
Much of the goal of these portals is to open up existing government data, but
[forms](https://data.wa.gov/Economics/Broadband-Project-Data-Entry/38rz-krmg?) provide a way for citizens to create data.
lets you enter data. A bunch of people have implemented them, but none seems to get accessed much.

```{r form_use_3}
forms <- sqldf('select portal, sum(viewCount) AS "total.view.count", count(*) AS "number.of.forms" FROM socrata where displayType = "form" group by portal')
ggplot(forms) +
  aes(x = number.of.forms, y = total.view.count, label = portal) + geom_text() +
```
I'm gonna remove opendata.socrata.com to make that easier to read.

```{r form_use_4}
forms <- sqldf('select portal, sum(viewCount) AS "total.view.count", count(*) AS "number.of.forms" FROM socrata where displayType = "form" and portal != "opendata.socrata.com" group by portal')
ggplot(forms) +
  aes(x = number.of.forms, y = total.view.count, label = portal) + geom_text() +
  scale_x_continuous('Number of forms in the portal') +
  scale_y_continuous('Total hits across all forms', labels = comma)
```

I hadn't seen [nmfs.socrata.com](https://nmfs.socrata.com) before.
It belongs to the [National Oceanic and Atmospheric Administration Fisheries Service](http://www.nmfs.noaa.gov),
which apparently used [a Socrata form](https://nmfs.socrata.com/Government/2011-Aquaculture-Public-Comments-Form/u5id-8nqp) to power a
[policy comments website](http://www.nmfs.noaa.gov/aquaculture/policy2/).

New York made a form for [feedback on the portal](https://data.ny.gov/dataset/Give-Feedback/fq3e-q75i?).

World Bank Open Finances made a
[form](https://finances.worldbank.org/dataset/Global-Open-Data-Calendar-Entry-Form/qdbh-rfd3?)
that populates an
[open data events calendar](https://finances.worldbank.org/dataset/Global-Open-Data-Calendar/g4sx-dwxc).


### Calendar
The [calendar](https://data.oregon.gov/dataset/Public-Meetings-3-Month-View-/4775-yg3b?)
is what it sounds like. I'm quite excited that Socrata presents calendars as data.
To me, anything could be turned into data, and data could be turned into
[anything](http://csvsoundsystem.com). But a lot of people don't realize this,
and I see calendars as one of those things that people don't see as data.

I'm also excited about Socrata calendars because people use them. A lot.

```{r calendar_use}
ggplot(subset(socrata.distinct, displayType == 'calendar')) + aes(x = viewCount) + geom_histogram() +
  scale_x_continuous('Hits per calendar', labels = comma) +
  scale_y_continuous('Frequency', labels = comma) +
  ggtitle("Someone likes her calendar.")
```

```{r calendar_use_2}
calendar.use <- subset(socrata, displayType == 'calendar')[c('portal', 'id', 'viewCount')]
calendar.use[order(calendar.use$viewCount, decreasing = T),][1:5,c('portal', 'id', 'viewCount')]
```

Check out Missouri's [open meetings calendar](https://data.mo.gov/Government-Administration/Open-Meetings-Calendar/mahp-izvx?).
They also have a different view of the calendar [outside the portal](http://www.mo.gov/meetings/),
with an RSS feed from the Socrata portal. But I don't know how the calendar on Socrata got so many views.

It looks like some other portals are using calendars a lot too,
but usually with several separate calendars instead of one huge one.

```{r calendar_use_3}
ggplot(sqldf('select portal, sum(viewCount) AS "total.view.count", count(*) AS "number.of.calendars" FROM "calendar.use" GROUP BY portal;')) +
  aes(x = number.of.calendars, y = total.view.count, label = portal) + geom_text() +
  scale_x_log10('Number of calendars in the portal', breaks = c(1,5,25)) +
  scale_y_log10('Total hits across all calendars', labels = comma, breaks = 10^(2:6))
```

Check out [Oregon's calendars](https://data.oregon.gov/browse?limitTo=calendars).

