---
title: Maps, charts and other possibilities
---


Socrata is more than a place to dump raw tables, or at least
it tries to be; you can make various charts and maps, and you
can serve non-tabular information to some degree. I'm wondering,
first, what all of these visualizations are and, second, which
ones people use where.

## Concepts

### Display types
There are many ways that a particular dataset could be visualized.
Socrata has `length(unique(socrata$displayType))` ways.
Here they are.

```{r display_types}
ggplot(sqldf('select displayType from socrata group by id')) + aes(x = displayType) + geom_bar() + coord_flip()
```

### View Types
Socrata is primarily used for storing tabular data,
but it can do a lot of other things too. It has four
[view](/!/socrata-genealogies#term-view)
types, one for each of four different internal
representations of the data.

* **tabular** is for tabular data, like a spreadsheet.
* **geo** is for map data. It's pretty similar to **tabular**,
    except that it has a special geospatial column type.
* **blobby** is for arbitrary files. You could put PDF files
    up if you want.
* **href** is for links to external files.

The most common view type is tabular.

```{r view_types}
ggplot(sqldf('select viewType from socrata group by id')) + aes(x = viewType) + geom_bar() + coord_flip()
```

### Those views with no view type
Above, you may have notice that lots of views have no view type.
At first, I thought they might just all be hrefs (external datasets).

```{r display_and_view_types}
ggplot(sqldf('select displayType, viewType from socrata group by id')) + aes(x = displayType, group = viewType, fill = viewType) + geom_bar() + coord_flip()                                                                                
```

Apparently not. Even though there's an href display type and a tabular
display type, we still find data with href and tabular view types and
empty display types. I looked at a few of these datasets to see what
was going on but didn't come up with any satisfactory answer.

So I don't really know why some views don't have view types.

## Popular view types
Let's focus on the things that are neither tables nor external links.

```{r not_boring}
not.boring <- subset(socrata.distinct, displayType != '' & displayType != 'table' & displayType != 'href' & displayType != 'blob')
not.boring$displayType <- factor(not.boring$displayType, levels = names(sort(table(not.boring$displayType), decreasing = T)))
ggplot(not.boring) + aes(x = displayType) + geom_bar() + coord_flip()
```

What are all of these types? Here are some examples.

* [intensitymap](https://opendata.socrata.com/dataset/Statewise-Prison-Population/dud8-q2qd?) (the only one)
* [calendar](https://data.oregon.gov/dataset/Public-Meetings-3-Month-View-/4775-yg3b?)
* [form](https://data.wa.gov/Economics/Broadband-Project-Data-Entry/38rz-krmg?)
* [fatrow](https://electionsdata.kingcounty.gov/dataset/Attorney-General-table-General-Election-2012/2y5z-ffa8?)
* [api](https://data.energystar.gov/Government/ENERGY-STAR-Certified-Geothermal-Heatpumps/2f83-p388?)
    (documentation is [here](https://data.energystar.gov/developers/docs/energy-star-certified-geothermal-heatpumps))

With so many types, I'm surprised that map and chart are pretty much the only ones that get used.
I can see why few people would make APIs, but you'd think people would want to try making pie charts
instead of the normal "chart".
