---
title: How up-to-date are data portal datasets?
tags: ['open-data']
---
```{r configure, echo = F}
source('15-dates.r')
opts_chunk$set(fig.width=10, fig.height=10, dpi = 42 * 5)
```

```{r up_to_date}
print(p1)
```

## Data portal metrics
According to A, B, C and D, getting data to the web, in whatever form
you can, is always better than nothing. How can you tell whether you're
doing a good job at this basic release of data? Data catalog software
typically provides one tool: the metric of number of datasets. We can go to
[data.gov.uk](https://data.gov.uk) and find out that it currently has
however many datasets, and we can go back next week and make sure that
the number is increasing. There are problems with this metric, which I
won't go into right now, but the metric does tell us something about how
much data we are releasing.

According to A, B, C and D, it is also important that we keep open data
up-to-date. Unfortunately, data catalog software don't currently
have great tools for helping you make sure that your data are indeed
up-to-date. Specifically, we don't have a number that we can look at
week-to-week to make sure that we are vaguely on track at keeping data
up-to-date. So I made one.

## Developing the metric

### Motivation for this



### Tuning
This metric involves a somewhat arbitrary cut-off, and that is somewhat
inconvenient; our updatedness value might change depending on whether
we use a cutoff of one year or half a year.

I was tempted to come up with a fancier metric that looked more at the
distribution of update times or chose a less-arbitrary cutoff for each
different portal, but that would be more complicated and might be harder
to understand. Instead of coming up with a less arbitrary cutoff, we
can check whether different values will give us the same value on our
updatedness metric. To say it fancy-like, we can check whether our
updatedness metric is robust to the cutoff.

A simple way of checking this is to plot the value of our updatedness
metric for different cutoff values. If we find a region where the value
of the updatedness metric doesn't change very much when we change the
cutoff, we could say that cutoffs in that region are safe to use.

## Updatedness by portal

### 



### Rivalries

Chicago v new york



