
Socrata is primarily used for storing tabular data,
but it can do a lot of other things too. It has four
[view](/!/socrata-genealogies#term-view)
types, one for each of four different internal
representations of the data.

* **tabular** is for tabular data, like a spreadsheet.
* **geo** is for map data. It's pretty similar to **tabular**,
    except that it has a special geospatial column type.
* **blobby** is for arbitrary files. You could put PDF files
    up if you want.
* **href** is for links to external files.

The most common view type is tabular.

```{r view_types}
ggplot(sqldf('select viewType from socrata group by id')) + aes(x = viewType) + geom_bar() + coord_flip()
```

### Those views with no view type
Above, you may have notice that lots of views have no view type.
At first, I thought they might just all be hrefs (external datasets).

```{r display_and_view_types}
ggplot(sqldf('select displayType, viewType from socrata group by id')) + aes(x = displayType, group = viewType, fill = viewType) + geom_bar() + coord_flip()                                                                                
```

Apparently not. Even though there's an href display type and a tabular
display type, we still find data with href and tabular view types and
empty display types. I looked at a few of these datasets to see what
was going on but didn't come up with any satisfactory answer.

So I don't really know why some views don't have view types.
