---
title: Geneology of dataset
---




Reference the ScraperWiki post on Twitter, the responses to me from Twitter, ...


## Metadata schema
Socrata documents how to search for [data within a dataset](https://opendata.socrata.com/api/docs/),
and it releases a bunch of examples [on GitHub](https://github.com/socrata). It also has a bunch of
cool-looking projects that seem to be parts of their infrastructure. (I'm referring to all of the
packages written in Scala.)

But I didn't find a great codebook/dictionary/schema for the dataset metadata.
([This](https://opendata.socrata.com/api/docs/views) is all I found.)
The genealogical information must be in there somewhere, so I'm gonna figure
it out and document it. And then I'll write about the geneology of datasets.

### `rowsUpdatedBy`?
```{r select.rowsUpdatedBy}
socrata.ancestry <- socrata[c('portal', 'id', 'rowsUpdatedBy')]
```

This field `rowsUpdatedBy` seemed to be present in sibling datasets.
Let's see where that takes us.

```{r common.ancestors}
sort(table(socrata.ancestry$rowsUpdatedBy), decreasing = T)[1:10]
```

`rowsUpdatedBy` is a user (4x4 Socrata id).

```{r top.ancestor}
top.ancestor <- names(sort(table(socrata.ancestry$rowsUpdatedBy), decreasing = T))[2]
print(top.ancestor) # 5fuc-pqz2
authors <- sort(table(subset(socrata, rowsUpdatedBy == top.ancestor)$tableAuthor.id), decreasing = T)[1:10]
```

Hmm that wasn't that interesting. Let's try something else

## Namespace for 4x4 ids
The 4x4 ids appear to be unique across portal. That is, no two different
datasets will have the same 4x4 id, even if they're in different portals.
On the other hand, if the same dataset is present in two portals, they
will have the same 4x4 id.


    diff data.hawaii.gov/views/5gah-bvex data.oregon.gov/views/5gah-bvex

```{r cross.portal.datasets}
cross.portal.datasets <- sort(table(socrata$id), decreasing = T)
print(cross.portal.datasets)[1:10]
```

Here's one of the top ones.

```{r top.cross.portal.dataset}
print(sort(table(socrata$id), decreasing = T)[1:10])
subset(socrata[c('portal', 'id', 'name')], id == '22u2-apnq')
```

In my download, the metadata files are exactly the same.

    diff explore.data.gov/views/22u2-apnq data.oregon.gov/views/22u2-apnq

The `oid` field tells us the same thing as the `id`.

```{r oid}
subset(socrata[c('portal', 'id', 'oid', 'name')], oid == 309331)
```

### Finding the original dataset
So how do we tell which portal is the original source?

There are a bunch of "owner" fields. I doubt that this owner is tied to a
portal, but it seems to work pretty well for identifying the source portal

```{r owner}
sort(table(socrata$owner.displayName), decreasing = T)[1:10]
```

Let's make sure that owner id and owner display name always match.

```{r owner.match}
owner.counts <- ddply(socrata[c('owner.id', 'owner.displayName')], c('owner.id', 'owner.displayName'), function(df) {c(n = nrow(df))})
max(owner.counts$n)
```


Same dataset in opendata.socrata.com and a different portal

```
                      portal        id    nrow ncol downloadCount
7790       dati.lombardia.it 3syc-54zf     234   56          1675
10425   opendata.socrata.com 54y8-wyde     234   56             9
```
